### 2012-10-21 paper_sean/methods.tex
minor changes with Sean

We evolved gaits first in simulation and then transferred them to the physical QuadraTot robot, as in Glette et al. 2012~\cite{glette}. 
Each run had a population size of 200 and lasted 200 generations. 20 runs were performed, each of which was identical except for the seed provided to the random number generator, which affected stochastic events such as mutation. 
Each gait was evaluated for 12 seconds \edit{JBY: how? As in fig 4 of \cite{yos:clune}, or another way?}, both in simulation and in reality. 
All HyperNEAT parameters were identical to those in Yosinski et al. 2011~\cite{yos:clune} except for the frequency of the sine wave input to the ANN, which was lowered from 0.0265Hz to 0.004Hz to decrease the number of servo shutdowns due to high-frequency gaits \edit{JBY: what? why?}.

@@

 %Sine frequency of 0.004 was used because we experimentally found that HyperNEAT was able to produce natural and high-performing gaits with low frequencies with this frequency. 
We also punished high-frequency gaits during evolving in simulation, which was effective at reducing the number of servo shutdowns observed when evolved gaits were transferred to the robot. 
If the average frequency of each leg \edit{JBY: any leg or all legs?} for the gait was higher than a threshold frequency (1.67 Hz), the fitness was punished by exponentially reducing the distance travelled by the robot:

>>>>

We evolved gaits first in simulation and then transferred them to the physical QuadraTot robot, as in Glette et al. 2012~\cite{glette}. 
Each run had a population size of 200 and lasted 200 generations. 20 runs were performed, each of which was identical except for the seed provided to the random number generator, which affected stochastic events such as mutation. 
Each gait was evaluated for 14 seconds in reality, with the first two
seconds being interpolated from a stationary position into the gait
and the last two seconds from the gait back into a stationary
position, respectively. This process was similar to that used in
\cite{yos:clune} and resulted in effectively 12 seconds of full speed
motion. In simulation gaits were simply evaluated for 12 seconds of
full speed motion. 
All HyperNEAT parameters were identical to those in Yosinski et al. 2011~\cite{yos:clune} except for the frequency of the sine wave input to the ANN, which was lowered from 4.24Hz to 0.640Hz. We lowered the sine wave frequency because after experimentation, we found that the previous settings for the sine wave resulted in high-frequency gaits that the servos could not run without shutting down.

@@

 %Sine frequency of 0.004 was used because we experimentally found that HyperNEAT was able to produce natural and high-performing gaits with low frequencies with this frequency. 
We also punished high-frequency gaits during evolving in simulation, which was effective at reducing the number of servo shutdowns observed when evolved gaits were transferred to the robot. We calculated the average frequency of the gait per leg by summing up the number of servo direction changes in all the legs throughout a gait and dividing by the length of the run and the number of legs. If the average frequency per leg for the gait was higher than a threshold frequency (1.67 Hz), the fitness was punished by exponentially reducing the distance travelled by the robot:

### 2012-10-21 paper_sean/results.tex
minor changes with Sean

% table!
\begin{table}
\caption{The number of evaluations per run, simulated (average) and real trial (max) velocities in [cm/s]. Data used from Yosinksi et al. 2011~\cite{yos:clune}, Glette et al. 2012~\cite{glette}, and Shen et al. 2012~\cite{haocheng}.
*17.8 cm/s was recorded on Glette et al. 2012~\cite{glette} but the same gait tried on our robot averaged over ten runs was 12.951 cm/s.}  %\tablabel{results}
\begin{center}
\begin{tabular}{|l|r|r|r|r|}
\hline
                                         & Evaluations  & Simulated Velocity  & Real Velocity \\
\hline
Parameterized gaits + optimization \cite{yos:clune}   &153    & --    & 4.6 \\
\hline
HyperNEAT in hardware \cite{yos:clune}                 & 180         & --         &   9.7     \\
\hline
RL PoWER Spline \cite{haocheng}                         & 300         & --         &   11.1 \\
\hline
GA + simulator \cite{glette}             & 60000       & 16.4       &   *13.0     \\
\hline
HyperNEAT + simulator [this paper]                     & 40000       & \textbf{25.4}       &   \textbf{14.5} \\
\hline
\end{tabular}
\end{center}
\label{resultsTable}
\end{table}

@@

% \begin{figure}
% \begin{center}
% \vspace{1cm}
% \epsfig{file=fit_gen.eps, width=8cm}
% \caption[ ]{Fitnesses of the champion gaits averaged over 20 runs, measured in distance travelled over 12 seconds [cm/s]. These gaits outperformed gaits evolved in simulation using Genetic Algorithm by 54.3\% in simulation and 5.4\% better in reality while outperforming gaits evolved using HyperNEAT in hardware by 49.5\%.}
% \end{center}
% \end{figure}
\figgp{avg_fit_vs_gen}{.48}{avg_freq_vs_gen}{.48}{Performance plots. \textbf{Left: }Fitnesses of the champion gaits averaged over 20 runs, measured in distance travelled over 12 seconds [cm/s]. These gaits outperformed gaits evolved in simulation using Genetic Algorithm by 54.3\% in simulation and 5.4\% better in reality while outperforming gaits evolved using HyperNEAT in hardware by 49.5\%. \edit{JBY: is it really standard deviation or standard error?}\textbf{Right: }Average frequencies of the gaits averaged over 20 runs. After implementing a policy which punished gaits with high frequencies, HyperNEAT was able to find high-performing, low frequency gaits. This policy was implemented to reduce servo shutdowns which heavily affected the reality performance of the robot.\edit{JBY: fig might be easier to read if y-axis was log scale?}}

>>>>

% table!
\begin{table}
\begin{center}
\begin{tabular}{|l|r|r|r|r|}
\hline
                                         & Evaluations  & Simulated Velocity  & Real Velocity \\
\hline
Parameterized gaits + optimization \cite{yos:clune}   &153    & --    & 5.8 \\
\hline
HyperNEAT in hardware \cite{yos:clune}                 & 180         & --         &   9.7     \\
\hline
RL PoWER Spline \cite{haocheng}                         & 300         & --         &   11.1 \\
\hline
GA + simulator \cite{glette}             & 60000       & 16.4       &   *13.0     \\
\hline
HyperNEAT + simulator [this paper]                     & 40000       & \textbf{25.4}       &   \textbf{14.5} \\
\hline
\end{tabular}
\end{center}
\label{resultsTable}
\caption{The number of evaluations per run, simulated (average) and real trial (max) velocities in [cm/s]. Data used from Yosinksi et al. 2011~\cite{yos:clune}, Glette et al. 2012~\cite{glette}, and Shen et al. 2012~\cite{haocheng}.
*17.8 cm/s was recorded on Glette et al. 2012~\cite{glette} but the same gait tried on our robot averaged over ten runs was 12.951 cm/s.}  %\tablabel{results}
\end{table}

@@

% \begin{figure}
% \begin{center}
% \vspace{1cm}
% \epsfig{file=fit_gen.eps, width=8cm}
% \caption[ ]{Fitnesses of the champion gaits averaged over 20 runs, measured in distance travelled over 12 seconds [cm/s]. These gaits outperformed gaits evolved in simulation using Genetic Algorithm by 54.3\% in simulation and 5.4\% better in reality while outperforming gaits evolved using HyperNEAT in hardware by 49.5\%.}
% \end{center}
% \end{figure}
\figgp{avg_fit_vs_gen}{.48}{avg_freq_vs_gen}{.48}{Performance plots. \textbf{Left: }Fitnesses of the champion gaits averaged over 20 runs, measured in distance travelled over 12 seconds [cm/s]. These gaits outperformed gaits evolved in simulation using Genetic Algorithm by 54.3\% in simulation and 5.4\% better in reality while outperforming gaits evolved using HyperNEAT in hardware by 49.5\%. \edit{Add Kyrre's runs, switch from STD to SE if necessary.}\textbf{Right: }Average frequencies of the gaits averaged over 20 runs. After implementing a policy which punished gaits with high frequencies, HyperNEAT was able to find high-performing, low frequency gaits. This policy was implemented to reduce servo shutdowns which heavily affected the reality performance of the robot. \edit{add line at 1.67 Hz}}

### 2012-10-21 paper_sean/robot.tex
minor changes with Sean

Each servo has a built-in safety mechanism that shuts itself off to prevent damage if the servo's current, range, temperature, or torque was too high~\cite{robotis}. During evolution, this safety mechanism frequently activated, and did so inconsistently, adding significant noise to the evaluation process. As pointed out in Yosinski et al. 2011~\cite{yos:clune}, \edit{is it better to say ``As pointed out in a previous study \cite{yos:clune}, gaits...''} gaits generated during evolution are highly variable and generate many shutdowns because they forced the servos to exert too much torque or command positions that are out of range. %JMC: Does requesting a value out of range really cause the servo to shut down? That seems like something we could easily fix (by never allowing an out-of-range value). Now that we do not allow values out of range, we still see lots of shutdowns...making me think that the problem was really torque. What do you think?
To prevent collisions between different pieces of the robot's body, we limited the allowable range of movement for the inner, outer, and hip joints to [-85\degree, +60\degree], [-113\degree, +39\degree], and [-28\degree, +28\degree], respectively. We also implemented the Smart Cropping System from Shen et al. 2012~\cite{haocheng}, which prevents combinations of joint positions for the inner and outer joint of each leg that generate extreme amounts of torque. A final method of reducing torque was to reduce the weight of the robot. Yosinski et al. 2011~\cite{yos:clune} had the small Linux computer that performed all computation on the robot, but we removed it and sent commands to the robot via a cable.

>>>>

Each servo has a built-in safety mechanism that shuts itself off to prevent damage if the servo's current, range, temperature, or torque was too high~\cite{robotis}. During evolution, this safety mechanism frequently activated, and did so inconsistently, adding significant noise to the evaluation process. As pointed out in a previous study~\cite{yos:clune}, gaits generated during evolution are highly variable and generate many shutdowns because they forced the servos to exert too much torque or command positions that are out of range. %JMC: Does requesting a value out of range really cause the servo to shut down? That seems like something we could easily fix (by never allowing an out-of-range value). Now that we do not allow values out of range, we still see lots of shutdowns...making me think that the problem was really torque. What do you think?
To prevent collisions between different pieces of the robot's body, we limited the allowable range of movement for the inner, outer, and hip joints to [-85\degree, +60\degree], [-113\degree, +39\degree], and [-28\degree, +28\degree], respectively. We also implemented the Smart Cropping System from Shen et al. 2012~\cite{haocheng}, which prevents combinations of joint positions for the inner and outer joint of each leg that generate extreme amounts of torque. A final method of reducing torque was to reduce the weight of the robot. Yosinski et al. 2011~\cite{yos:clune} had the small Linux computer that performed all computation on the robot, but we removed it and sent commands to the robot via a cable.
